A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-02-07 paper.tex
spell checked and updated schematic

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
    \caption[]{
      \label{fig:software}
      The various objects and their relation to one another in the \code{aneris}
      code base as well as a short description of their scope.  
    }
  \end{center}
\end{figure}

@@

In order to help identify cases where overrides may be needed,
harmonization \textit{diagnostics} are provided which report the relative
difference between harmonized and unharmonized trajectories at their mid and
end-points when these values exceed specified threshold. The diagnostic reporting thresholds  are configurable by the user, but defaults of $400$\%
and $200$\%, respectively, are provided based on experiences of the authors'
use of \code{aneris} to date.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization results for two scenarios from
the \gls{ssp} scenario library\footnote{We refer the reader to the broad
  literature discussing the SSPs, e.g.,
  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
  depth discussion of the scenario architecture and design.} are presented here. We use the
SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
(referred to as SSP2-Ref) as an example because MESSAGE-GLOBIOM is
the marker scenario\footnote{The ``marker'' scenario concept is used to
  designate the archetype scenario used as a reference within each SSP scenario family. See
  \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
  description.} for this \gls{ssp}. This SSP2 scenario lies between two
\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing in this concept is the energy inbalance at the top of the atmsophere caused by anthropogenic influences relative to a pre-industrial reference point. Higher radiative forcing leads to larger global changes, such as surface temperture. We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
  for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
additionally present the results for the SSP2-based mitigation scenario leading
to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
scenario is chosen because mitigation technologies and policies are enacted
causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
negative \cotwo emissions in some regions and sectors due to carbon capture and
sequestration and afforestation. A scenario in which negative emissions play a
role in mitigation strategies is particularly important because of the
sensitivity of key indicators, such as end-of-century radiative forcing (which
is used to estimate mean global temperature response), to the timing and
magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
scenarios represent two contrasting cases in the use of a harmonization approach
and thus provide a case study as to its general applicability. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
aggregate \gls{ghg} emissions, in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions, and harmonization is performed at the native regional resolution of each model. Historical
data is taken from previously described \gls{luc} and anthropogenic sources,
which comprise 10 separate pollutant and \gls{ghg} species and 12 sectors shown
in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
each scenario, and therefore 1940 trajectories were harmonized in total for these two illustrative scenarios. 
\noxx
generated from the Energy sector provides an example of an emissions species and
sector in which all regions were satisfactorily harmonized with the default
methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
Table \ref{tab:nox} describes the parameters that underlie the choice of method
for each harmonized trajectory. Importantly, the default methods provide
regional trajectories for both the reference and mitigation scenarios that
match the general unharmonized model behavior and also result in global aggregated emissions
that are similar to the unharmonized trajectory.

@@

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized (without markers) and unharmonized (with
      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
      lines) scenarios. The left column shows the deault harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
      while panels \textbf{b} and \textbf{c} show  trajectories for the
      \gls{afr} region and (total) global emissions, respectively. Notably, the
      \gls{sas} and \gls{afr} regional trajectories are distorted
      when the default methods are used due to the harmonization-year difference between history and
      model results in both scenarios. The distortion is large enough to affect
      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistantcy with the unharmonized model scenario at the regional and global levels. }
  \end{center}
\end{figure}

>>>>

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
    \caption[]{
      \label{fig:software}
      The various objects and their relation to one another in the \code{aneris}
      code base as well as a short description of their scope of concern.  
    }
  \end{center}
\end{figure}

@@

In order to help identify cases where overrides may be needed, harmonization
\textit{diagnostics} are provided which report the relative difference between
harmonized and unharmonized trajectories at their mid and end-points when these
values exceed specified thresholds. The diagnostic reporting thresholds are
configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
provided based on experiences of the authors' use of \code{aneris} to date.

@@

In order to show a representative cross section of the performance of the
\code{aneris} harmonization procedure, we focus on the harmonization of results
of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization
results for two scenarios from the \gls{ssp} scenario library\footnote{We refer
  the reader to the broad literature discussing the SSPs, e.g.,
  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
  depth discussion of the scenario architecture and design.} are presented
here. We use the SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of
the road'', scenario (referred to as SSP2-Ref) as an example because
MESSAGE-GLOBIOM is the marker scenario\footnote{The ``marker'' scenario concept
  is used to designate the archetype scenario used as a reference within each
  SSP scenario family. See \cite{vuuren_new_2013} and \cite{Riahi2017153} for a
  more lengthy description.} for this \gls{ssp}. This SSP2 scenario lies between
two \glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing
  in this context is the energy imbalance at the top of the atmosphere caused by
  anthropogenic influences relative to a pre-industrial reference point. Higher
  radiative forcing leads to larger global changes, such as surface
  temperature. We refer the reader to \cite{IPCCWG1PhysicalStocker2013} for a
  more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
additionally present the results for the SSP2-based mitigation scenario leading
to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
scenario is chosen because mitigation technologies and policies are enacted
causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
negative \cotwo emissions in some regions and sectors due to carbon capture and
sequestration and afforestation. A scenario in which negative emissions play a
role in mitigation strategies is particularly important because of the
sensitivity of key indicators, such as end-of-century radiative forcing (which
is used to estimate mean global temperature response), to the timing and
magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
scenarios represent two contrasting cases in the use of a harmonization approach
and thus provide a case study as to its general applicability. Figure
\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
aggregate \gls{ghg} emissions, in each scenarios.

@@

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the RCPs (Figure
\ref{fig:regions}); harmonization is performed directly on the native regional
spatial resolution. Historical data is taken from previously described \gls{luc}
and anthropogenic sources, which comprise 10 separate pollutant and \gls{ghg}
species and 12 sectors shown in Table \ref{tab:sp}. A total of 970 distinct
trajectories were harmonized for each scenario, and therefore 1940 trajectories
were harmonized in total for these two illustrative scenarios.  \noxx generated
from the Energy sector provides an example of an emissions species and sector in
which all regions were satisfactorily harmonized with the default
methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
Table \ref{tab:nox} describes the parameters that underlie the choice of method
for each harmonized trajectory. Importantly, the default methods provide
regional trajectories for both the reference and mitigation scenarios that match
the general unharmonized model behavior and also result in global aggregated
emissions that are similar to the unharmonized trajectory.

@@

\begin{figure}
  \begin{center}
    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
    \caption[]{
      \label{fig:co}
      CO Industrial Sector harmonized (without markers) and unharmonized (with
      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
      lines) scenarios. The left column shows the default harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
      while panels \textbf{b} and \textbf{c} show  trajectories for the
      \gls{afr} region and (total) global emissions, respectively. Notably, the
      \gls{sas} and \gls{afr} regional trajectories are distorted
      when the default methods are used due to the harmonization-year difference between history and
      model results in both scenarios. The distortion is large enough to affect
      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistency with the unharmonized model scenario at the regional and global levels. }
  \end{center}
\end{figure}
